An Embedding Framework for the Design and Analysis of Consistent Polyhedral Surrogates

### 2024-02-22 poly-embed-journal/poly-embed-journal-camera-ready.tex
camera-ready tweaks
Adding in TikZ for Figures 1 and 6

\tikzstyle{cell}=[dashed,thick]
\tikzstyle{simplex}=[thick]
\newcommand{\tikzfigscale}{2.5}
\newcommand{\placefiglabel}[1]{\node at (1.1,0,1.5) {#1};}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

@@

Intuitively, $\Gamma(p)$ is the set of reports which optimize expected loss under a given distribution $p$, and $\Gamma_r$ is the set of distributions for which the report $r$ optimizes the expected loss.
For example, the \emph{mode} is the %finite
property $\mode(p) = \argmax_{y\in\Y} p_y$, and captures the set of optimal reports for 0-1 loss: for each distribution over the labels, one should report the most likely label.
In this case we say 0-1 loss (directly) \emph{elicits} the mode, as we formalize below.
% This terminology comes from the information elicitation \jessiet{information vs property... do we care?}\raft{we could also say economics literature; not ``property'' though since that doesn't offer an explanation of the word ``elicits''} literature~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting}, in which a report $r$ is elicited from some forecaster by scoring her with a loss on the observed outcome $y$.

@@

When working with a restricted hypothesis class $\H$, as opposed to the set of all measurable functions in Definition~\ref{def:consistency}, then the corresponding notion of consistency (called \emph{$\H$-consistency}) is no longer equivalent to calibration~\citep{long2013consistency,kuznetov2014multi,zhang2020convex};
see \S~\ref{sec:conclusion} for further discussion.

@@

\begin{figure}
	\begin{minipage}{0.3\linewidth}
	\centering
	\includegraphics[width=0.95\linewidth]{figs/0-1-br.pdf}
%		\caption{Bayes risk of the 0-1 loss.}
%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\linewidth}
	\centering		\includegraphics[width=0.95\linewidth]{figs/hinge-br.pdf}
%		\caption{Bayes risk of hinge loss.}
%		\label{fig:hinge-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\linewidth}
	\centering
	\includegraphics[width=0.95\linewidth]{figs/logistic-br.pdf}
%	\caption{Bayes risk of the logistic loss.}
%	\label{fig:logistic-br}
\end{minipage}
\caption{Bayes risks $\risk L : p \mapsto \inf_u \inprod{p}{L(u)}$ of 0-1, hinge, and logistic losses, respectively, plotted as a function of $p_1 = \Pr[Y=1]$.
	Observe that the Bayes risks of 0-1 and hinge loss are both piecewise linear and concave, while the Bayes risk of logistic loss is not piecewise linear.  Proposition~\ref{prop:embed-bayes-risks} states that embedding is equivalent to matching Bayes risks, confirming that hinge loss (M) embeds twice 0-1 loss (L), but logistic loss (R) does not.}
\label{fig:bayes-risks-01}
\end{figure}

@@

\begin{figure}
\begin{center}
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{tikz/abstain-link-linf.pdf}
\end{minipage}\hfill
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{tikz/abstain-link-U-regions-l1.pdf}
\end{minipage}\hfill
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{tikz/abstain-link-l1.pdf}
\end{minipage}\hfill
\caption{Designing links for $\BEP$ with $d=2$ using Construction~\ref{const:eps-thick-link}. The embedding is shown in bold labeled by the corresponding reports. (L) The link envelope $\Psi$ resulting from Construction~\ref{const:eps-thick-link} using $\|\cdot\|_\infty$ and $\epsilon = 1/2$, and a possible link $\psi$ which matches eq.~\eqref{eq:abstain-link} from~\cite{ramaswamy2018consistent}.  (M) An illustration of the thickened sets for two sets $U, U' \in \U$, using $\|\cdot\|_1$ and $\epsilon = 1$. (R) The envelope $\Psi$ and link $\psi$ %from Construction~\ref{const:eps-thick-link}
  using $\|\cdot\|_1$ and $\epsilon = 1$.}
\label{fig:abstain-links}
\end{center}
\end{figure}

@@

@@

\begin{figure}
	\begin{center}
		\begin{minipage}{0.32\linewidth}
			\includegraphics[width=\linewidth]{tikz/lovasz-link-psi-star.pdf}
		\end{minipage}\hfill
		\begin{minipage}{0.32\linewidth}
			\includegraphics[width=\linewidth]{tikz/lovasz-link-psi-diamond.pdf}
		\end{minipage}\hfill
		\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{tikz/lovasz-link-envelope.pdf}
		\end{minipage}\hfill
		\caption{Links $\psi^*$ and $\psi^\diamond$ such that $(L^f, \psi^*)$ and $\psi^\diamond$ are calibrated with respect to $\ellabs$ for all suitable $f$. Points in each region link to the embedding point contained in the region. Both are constructed via the link envelope $\Psi$ from Construction~\ref{const:eps-thick-link}, which yields possible choices for calibrated links.
			}
			%Points in the dark green regions link to $\vec 0$; points in light green link to a permutation of $(0,\pm 1)$, and points in yellow regions link to $(\pm 1, \pm 1)$.\jessie{Not sure how to phrase this?  Would we just be better off explicitly labeling things?}\raf{Agreed that it's a little unclear, but not sure labeling would be better... let's discuss next call}}
		\label{fig:lovasz-links}
	\end{center}
\end{figure}

@@

\subsection{Surrogates for top-$k$ classification}
\label{sec:top-k}
% In settings like object recognition and information retrieval, the top-$k$ classification problem arises in which one predicts a set $S$ of size $k$, and upon realizing outcome $y$, receives loss $\elltopk(S)_y = \ones\{y\notin S\}$~\citep{lapin2015top, lapin2016loss, lapin2018analysis,yang2018consistency,berrada2018smooth,rastegari2011scalable,reddi2019stochastic}.
% Many of the existing surrogates for top-$k$ classification consider the following three desiderata: convexity, calibration, and piecewise linear (``hinge-like'') structure.
% Satisfying all three desiderata simultaneously had not been observed until the work of~\citet{finocchiaro2022consistenttopk}.\raft{Not sure you can use ``observed'' like that (came up a few times)}
% The surrogates that are convex and calibrated but not polyhedral often elicit the entire distribution $p \in \simplex$, and take the link to be the top $k$ elements of $p$.
% \citet{lapin2016loss} observes that this approach learns more information than desired and often requires more samples
% % for surrogate regret to approach target regret
% ~\citep{frongillo2021surrogate}.\jessiet{right?}
% \citet{yang2018consistency} propose a piecewise linear surrogate that is calibrated for top-$k$ classification, but is not convex, generally making optimization more difficult.
% Now consider the class of polyhedral surrogates whose calibration for top-$k$ prediction is unknown~\citep{yang2018consistency,lapin2016loss,lapin2018analysis}: Theorem~\ref{thm:poly-embeds-discrete} says that every polyhedral surrogate embeds some discrete loss, and Theorem~\ref{thm:embed-poly-main} states that embedding implies calibration by some link for this discrete loss; thus, we can use embedding to analyze the (non)calibration of existing polyhedral surrogates with respect to top-$k$ by comparing the embedded losses to top-$k$.
In settings like object recognition and information retrieval, it is natural to predict a set $S$ of labels.
In top-$k$ classification, one requires $|S|=k$, and given the true label $y$, the target loss is $\elltopk(S)_y = \ones\{y\notin S\}$~\citep{lapin2015top, lapin2016loss, lapin2018analysis,yang2018consistency,berrada2018smooth,rastegari2011scalable,reddi2019stochastic}.
In the literature on surrogates for top-$k$ classification, one goal has been to find a surrogate satisfying the following three desiderata: convexity, consistency, and piecewise linear (``hinge-like'') structure.
\citet{yang2018consistency} show that a number of previously proposed polyhedral losses, i.e., those which are convex and hinge-like, are inconsistent.
They further conjecture that perhaps no surrogate could satisfy all three properties.
%\raft{Also, I cut the smooth vs hinge-like discussion since we don't really need the full motivation -- saying ``these weirdos wanted to do XYZ; who knows why'' is good enough, since our point is that our framework is useful for whatever people are trying to do \jessie{makes sense! \#burn}}

@@

\begin{table*}
	\centering
	\begin{tabular}{ccccc}
		& $\prop{\Li{2}}$ & $\prop{\Li{3}}$ & $\prop{\Li{4}}$ & $\prop{L^k}$\\
		\hline \hline
		\rotatebox[origin=c]{90}{$k = 2$\hspace*{-2.5cm}} & \input{tikz/n4-k2-psi2-slice} & \input{tikz/n4-k2-psi3-slice} & \input{tikz/n4-k2-psi4-slice} & \input{tikz/n4-k2-psi6-slice}\\ 
		\rotatebox[origin=c]{90}{$k = 3$\hspace*{-2.5cm}} & \input{tikz/n4-k3-psi2-slice} & \input{tikz/n4-k3-psi3-slice} & \input{tikz/n4-k3-psi4-slice} & \input{tikz/n4-k3-psi6-slice}\\ 
	\end{tabular}
	\caption{
		Visualizations of the properties elicited by the losses (embedded by) $\Li{2}$, $\Li{3}$, $\Li{4}$ studied by \citet{yang2018consistency} and \citet{finocchiaro2022consistenttopk}, and $L^k$ in eq.~\eqref{eq:topk-embedding}.
    We take $n=4$ and $k \in \{2,3\}$, and visualize in 2 dimensions by fixing $p_4 = 1/4$. 
		%The dashed blue lines form cells whose elements are distributions $p$ corresponding to the same $u \in \prop{\elli k}(p)$ labeling the cell. 
		%As the link $\psi$ must be deterministic, in order for the $(\Li{k}, \psi)$ to be consistent with respect to $\elli k$, each cell outlined in black is fully contained in exactly one cell from the dashed blue lines. 
		The blue-filled regions are cells of the surrogate property which cross the dashed blue lines of the target property, exhibiting inconsistency (see Figure~\ref{fig:abstain-intuition-inconsistent}).
    Intuitively, the inconsistency arises from ambiguity in the top $k$ elements of the optimal surrogate report.  
		% See the caption for Figure~\ref{fig:ordered-partition} on how to interpret what the colored regions say about inconsistency with respect to top-$k$ classification.
	}
	\label{tab:loss-slices}
\end{table*}

@@

\begin{figure}[t]
	\begin{minipage}{0.47\linewidth}
		\centering
		\includegraphics[width=0.95\linewidth]{figs/modhinge.pdf}
		%		\caption{Bayes risk of the 0-1 loss.}
		%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centering		\includegraphics[width=0.95\linewidth]{figs/modhinge-br.pdf}
		%		\caption{Bayes risk of hinge loss.}
		%		\label{fig:hinge-br}
	\end{minipage}
	\caption{(L) Expected modified hinge loss for fixed distribution; (R) Bayes risk of modified hinge still matches the Bayes risk of hinge.}
	\label{fig:modified-hinge}
\end{figure}

>>>>

\tikzstyle{cell}=[dashed,thick]
\tikzstyle{simplex}=[thick]
\newcommand{\tikzfigscale}{2.5}
\newcommand{\tikzfigscalepgf}{.45}
\newcommand{\placefiglabel}[1]{\node at (1.1,0,1.5) {#1};}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

@@

Intuitively, $\Gamma(p)$ is the set of reports which optimize expected loss under a given distribution $p$, and $\Gamma_r$ is the set of distributions for which the report $r$ optimizes the expected loss.
For example, the \emph{mode} is the %finite
property $\mode(p) = \argmax_{y\in\Y} p_y$, and captures the set of optimal reports for 0-1 loss: for each distribution over the labels, one should report the most likely label.
In this case we say 0-1 loss (directly) \emph{elicits} the mode, as we formalize below.
% This terminology comes from the information elicitation \jessiet{information vs property... do we care}\\
\raft{we could also say economics literature; not ``property'' though since that doesn't offer an explanation of the word ``elicits''} literature~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting}, in which a report $r$ is elicited from some forecaster by scoring her with a loss on the observed outcome $y$.

@@

When working with a restricted hypothesis class $\H$, as opposed to the set of all measurable functions in Definition~\ref{def:consistency}, then the corresponding notion of consistency (called \emph{$\H$-consistency}) is no longer equivalent to calibration~\citep{long2013consistency,zhang2020convex,kuznetsov2014multi};
see \S~\ref{sec:conclusion} for further discussion.

@@

\begin{figure}
	\begin{minipage}{0.3\linewidth}
	\centering
	\input{tikz/BR01.tex}
%		\caption{Bayes risk of the 0-1 loss.}
%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\linewidth}
	\centering		\input{tikz/BRhinge.tex}
%		\caption{Bayes risk of hinge loss.}
%		\label{fig:hinge-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\linewidth}
	\centering
	\input{tikz/BRlogistic.tex}
%	\includegraphics[width=0.95\linewidth]{figs/logistic-br.pdf}
%	\caption{Bayes risk of the logistic loss.}
%	\label{fig:logistic-br}
\end{minipage}
\caption{Bayes risks $\risk L : p \mapsto \inf_u \inprod{p}{L(u)}$ of 0-1, hinge, and logistic losses, respectively, plotted as a function of $p_1 = \Pr[Y=1]$.
	Observe that the Bayes risks of 0-1 and hinge loss are both piecewise linear and concave, while the Bayes risk of logistic loss is not piecewise linear.  Proposition~\ref{prop:embed-bayes-risks} states that embedding is equivalent to matching Bayes risks, confirming that hinge loss (M) embeds twice 0-1 loss (L), but logistic loss (R) does not.}
\label{fig:bayes-risks-01}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{tikz/abstain-link-linf.pdf}
\end{minipage}\hfill
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{tikz/abstain-link-U-regions-l1.pdf}
\end{minipage}\hfill
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{tikz/abstain-link-l1.pdf}
\end{minipage}\hfill
\caption{Designing links for $\BEP$ with $d=2$ using Construction~\ref{const:eps-thick-link}. The embedding is shown in bold labeled by the corresponding reports. (L) The link envelope $\Psi$ resulting from Construction~\ref{const:eps-thick-link} using $\|\cdot\|_\infty$ and $\epsilon = 1/2$, and a possible link $\psi$ which matches eq.~\eqref{eq:abstain-link} from~\cite{ramaswamy2018consistent}.  (M) An illustration of the thickened sets for two sets $U, U' \in \U$, using $\|\cdot\|_1$ and $\epsilon = 1$. (R) The envelope $\Psi$ and link $\psi$ %from Construction~\ref{const:eps-thick-link}
  using $\|\cdot\|_1$ and $\epsilon = 1$.}
\label{fig:abstain-links}
\end{center}
\end{figure}

@@

\begin{figure}[t]
	\begin{center}
		\begin{minipage}{0.32\linewidth}
			\includegraphics[width=\linewidth]{tikz/lovasz-link-psi-star.pdf}
		\end{minipage}\hfill
		\begin{minipage}{0.32\linewidth}
			\includegraphics[width=\linewidth]{tikz/lovasz-link-psi-diamond.pdf}
		\end{minipage}\hfill
		\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{tikz/lovasz-link-envelope.pdf}
		\end{minipage}\hfill
		\caption{Links $\psi^*$ and $\psi^\diamond$ such that $(L^f, \psi^*)$ and $\psi^\diamond$ are calibrated with respect to $\ellabs$ for all suitable $f$. Points in each region link to the embedding point contained in the region. Both are constructed via the link envelope $\Psi$ from Construction~\ref{const:eps-thick-link}, which yields possible choices for calibrated links.
			}
			%Points in the dark green regions link to $\vec 0$; points in light green link to a permutation of $(0,\pm 1)$, and points in yellow regions link to $(\pm 1, \pm 1)$.\jessie{Not sure how to phrase this?  Would we just be better off explicitly labeling things?}\raf{Agreed that it's a little unclear, but not sure labeling would be better... let's discuss next call}}
		\label{fig:lovasz-links}
	\end{center}
\end{figure}

@@

@@

\subsection{Surrogates for top-$k$ classification}
\label{sec:top-k}
% In settings like object recognition and information retrieval, the top-$k$ classification problem arises in which one predicts a set $S$ of size $k$, and upon realizing outcome $y$, receives loss $\elltopk(S)_y = \ones\{y\notin S\}$~\citep{lapin2015top, lapin2016loss, lapin2018analysis,yang2018consistency,berrada2018smooth,rastegari2011scalable,reddi2019stochastic}.
% Many of the existing surrogates for top-$k$ classification consider the following three desiderata: convexity, calibration, and piecewise linear (``hinge-like'') structure.
% Satisfying all three desiderata simultaneously had not been observed until the work of~\citet{finocchiaro2022consistenttopk}.\raft{Not sure you can use ``observed'' like that (came up a few times)}
% The surrogates that are convex and calibrated but not polyhedral often elicit the entire distribution $p \in \simplex$, and take the link to be the top $k$ elements of $p$.
% \citet{lapin2016loss} observes that this approach learns more information than desired and often requires more samples
% % for surrogate regret to approach target regret
% ~\citep{frongillo2021surrogate}.\jessiet{right?}
% \citet{yang2018consistency} propose a piecewise linear surrogate that is calibrated for top-$k$ classification, but is not convex, generally making optimization more difficult.
% Now consider the class of polyhedral surrogates whose calibration for top-$k$ prediction is unknown~\citep{yang2018consistency,lapin2016loss,lapin2018analysis}: Theorem~\ref{thm:poly-embeds-discrete} says that every polyhedral surrogate embeds some discrete loss, and Theorem~\ref{thm:embed-poly-main} states that embedding implies calibration by some link for this discrete loss; thus, we can use embedding to analyze the (non)calibration of existing polyhedral surrogates with respect to top-$k$ by comparing the embedded losses to top-$k$.

\begin{table*}[t]
	\centering
	\begin{tabular}{cc@{}c@{}c@{}c}
		& $\prop{\Li{2}}$ & $\prop{\Li{3}}$ & $\prop{\Li{4}}$ & $\prop{L^k}$\\[4pt]
		% \hline
		\rotatebox[origin=c]{90}{$k = 2$\hspace*{-2.5cm}} & \input{tikz/n4-k2-psi2-slice} & \input{tikz/n4-k2-psi3-slice} & \input{tikz/n4-k2-psi4-slice} & \input{tikz/n4-k2-psi6-slice}\\ 
		\rotatebox[origin=c]{90}{$k = 3$\hspace*{-2.5cm}} & \input{tikz/n4-k3-psi2-slice} & \input{tikz/n4-k3-psi3-slice} & \input{tikz/n4-k3-psi4-slice} & \input{tikz/n4-k3-psi6-slice}\\
     % \hline
	\end{tabular}
	\caption{
		Visualizations of the properties elicited by the losses (embedded by) $\Li{2}$, $\Li{3}$, $\Li{4}$ studied by \citet{yang2018consistency} and \citet{finocchiaro2022consistenttopk}, and $L^k$ in eq.~\eqref{eq:topk-embedding}.
    We take $n=4$ and $k \in \{2,3\}$, and visualize in 2 dimensions by fixing $p_4 = 1/4$. 
		%The dashed blue lines form cells whose elements are distributions $p$ corresponding to the same $u \in \prop{\elli k}(p)$ labeling the cell. 
		%As the link $\psi$ must be deterministic, in order for the $(\Li{k}, \psi)$ to be consistent with respect to $\elli k$, each cell outlined in black is fully contained in exactly one cell from the dashed blue lines. 
		The blue-filled regions are cells of the surrogate property which cross the dashed blue lines of the target property, exhibiting inconsistency (see Figure~\ref{fig:abstain-intuition-inconsistent}).
    Intuitively, the inconsistency arises from ambiguity in the top $k$ elements of the optimal surrogate report.  
		% See the caption for Figure~\ref{fig:ordered-partition} on how to interpret what the colored regions say about inconsistency with respect to top-$k$ classification.
	}
	\label{tab:loss-slices}
\end{table*}

In settings like object recognition and information retrieval, it is natural to predict a set $S$ of labels.
In top-$k$ classification, one requires $|S|=k$, and given the true label $y$, the target loss is $\elltopk(S)_y = \ones\{y\notin S\}$~\citep{lapin2015top, lapin2016loss, lapin2018analysis,yang2018consistency,berrada2018smooth,rastegari2011scalable,reddi2019stochastic}.
In the literature on surrogates for top-$k$ classification, one goal has been to find a surrogate satisfying the following three desiderata: convexity, consistency, and piecewise linear (``hinge-like'') structure.
\citet{yang2018consistency} show that a number of previously proposed polyhedral losses, i.e., those which are convex and hinge-like, are inconsistent.
They further conjecture that perhaps no surrogate could satisfy all three properties.
%\raft{Also, I cut the smooth vs hinge-like discussion since we don't really need the full motivation -- saying ``these weirdos wanted to do XYZ; who knows why'' is good enough, since our point is that our framework is useful for whatever people are trying to do \jessie{makes sense! \#burn}}

@@

@@

\begin{figure}[t]
	\begin{minipage}{0.47\linewidth}
		\centering
		\input{tikz/mod_hinge.tex}		
		%\includegraphics[width=0.95\linewidth]{figs/modhinge.pdf}
		%		\caption{Bayes risk of the 0-1 loss.}
		%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\linewidth}
		\centering		
		\input{tikz/BRmod_hinge.tex}		
		%\includegraphics[width=0.95\linewidth]{figs/modhinge-br.pdf}
		%		\caption{Bayes risk of hinge loss.}
		%		\label{fig:hinge-br}
	\end{minipage}
	\caption{(L) Expected modified hinge loss for fixed distribution; (R) Bayes risk of modified hinge still matches the Bayes risk of hinge.}
	\label{fig:modified-hinge}
\end{figure}
